### 2021-07-15 chapters/2_introduction.tex
feedback processing

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering the mutants\cite{Ma2016,Yu2019PossibilityScope}.
Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
\newline
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated \todo{make the goal clearer: no requirements on which mutants get clustered: aim at generic mutant clustering}.
We will do this with two different approaches: qualitative and quantitative. 
We structure our research as follows:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement a clustering algorithm that will cluster one ore more characteristics of the mutants. 
We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.

>>>>

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering the mutants\cite{Ma2016,Yu2019PossibilityScope}.
Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
\newline
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated within the Java programming language.
This research aims to remove the scoping limitations present in existing research.
There should be no requirements for which mutants can be clustered.
We will do this with two different approaches: qualitative and quantitative. 
We structure our research as follows:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement a clustering algorithm that will cluster one ore more characteristics of the mutants. 
We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.

### 2021-07-15 chapters/3_background.tex
feedback processing

\chapter{Background}
\label{ch:background}
This chapter presents the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
Next, we will compare existing mutation testing tools. We start of by reviewing existing literature and then extend the comparison. 
This thesis relies on clustering techniques.
We review the state of the art on clustering algorithms and secondly the state of the art on machine learning models specialized in clustering.

\section{Terminology}
Subsuming mutants.
Mutation testing.
Levenshtein distance.
Maven.
Gradle.
Mutator.
test engine.
weighted mutation score.
\section{Project selection}
\label{ch:project_selection}
We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
\newline
We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
The unfiltered sample contains 50+ projects.
From there we filter out all the projects that contain failing tests and that are not libraries or applications.
The result consists of a sample with fifteen projects in total.

>>>>

\chapter{Background}
\label{ch:background}
This chapter presents the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
This thesis relies on clustering techniques.
We review the state of the art on clustering algorithms and secondly the state of the art on machine learning models specialized in clustering.

\section{Terminology}
\textbf{Mutation testing}
a form of white box testing in which testers change specific components of an application's source code to ensure a software test suite will be able to detect the changes.
\textbf{Levenshtein distance} the minimum number of single-character edits (insertions, deletions or substitutions) required to change one piece of text into the other.
\textbf{Maven} build tooling for Java software projects.
\textbf{Gradle} build tooling for Java and Android software projects.
\textbf{Mutator} definition of how the original source code is transformed into a mutant.
\textbf{test engine} a test engine facilitates discovery and execution of tests. Examples for Java are TestNG and JUnit.
\textbf{weighted mutation score} mutation score calculation designed to be comparable with the mutation score of a full set(see Chapter \ref{ch:weighted_score} for a detailed explanation).

\section{Project selection}
\label{ch:project_selection}
We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
\newline
We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
The unfiltered sample contains 50+ projects.
From there we filter out all the projects that contain failing tests and that are not libraries or applications.
The result consists of a sample with fifteen projects in total.

### 2021-07-15 chapters/4_mutation_tool.tex
feedback processing

\chapter{Mutation testing tools}
\label{ch:mutation-tool}

Before we can start clustering mutants we need to select a tool that helps us generate mutants. 
Our goal is to find the tool that generates the most amount of mutants. 
\newline
First, we review existing literature. 
Secondly we extend the comparison by adding information about the current state of the tools.
With the combination of the information gathered we make a decision on what tool to use for this thesis.

\section{State of the art documented}
\label{ch:tooling-literature}
The literature used several criteria for comparing mutation testing tools for Java:
\begin{itemize}
  \item Execution time of each tool.
  \item The cost as the number of test cases that needs to be generated and the the number of equivalent mutants that would have to be inspected.
  \item Effectiveness of the mutation adequate test suite of each tool. Every test of the suite adds to the effectiveness score. If a test is removed less than 100\% effectiveness can be achieved. To evaluate the effectiveness of each tool's mutation adequate test suite, a cross-testing technique is applied. The adequate test suite per tool is run on the set of mutants generated by the other tools.
\end{itemize}
Kintis et al. analysed and compared the performance of PIT,  MuJava and Major\cite{Kintis2016AnalysingStudy}. 
They started out with performing a mini literature survey on mutation testing tools used in existing research on test effectiveness in Java. 
With a data set of twelve methods and six Java projects they evaluated each tool using a cross-testing technique.
As a result they found that the most effective mutation adequate test suite was from MuJava, followed by Major and PIT.
The score of the application cost was the inverse of the previous ranking: PIT generated the smallest set of equivalent mutants and required the least test cases.
During their research they also compared the different mutant operators the tools supported. 
An overview of mutant operators(also known as mutators) per tool was created. 
They compared the mutant operators and reported overlapping operators.
\newline
Marki and Lindstrom performed their research on the same three mutation testing tools.\cite{Marki2017MutationJava}. 
The same cross-testing technique used by Kintis et al., was applied for three small Java programs. 
These programs are popular programs used in testing literature.
They found that the mutation tools do not subsume each other. 
According to the research a ranking of the tools would be as follows; MuJava generated the strongest mutants followed by Major and PIT.
Furthermore they found that MuJava generated significantly more equivalent mutants\cite{Marki2017MutationJava}.
MuJava also had an execution time twice the amount of that of Major and PIT combined.
\newline
Laurent et al., introduced PIT+, a version of PIT with extended set of mutators\cite{Laurent2017AssessingPIT}.
They used the same tests suites that were generated by kintis et al., and combined these into an adequate test suite that would detect the combined set of mutants generated by PIT, MuJava and Major. 
They discovered that the set of mutants generated by PIT+ was stronger than the combined sets generated by the other three tools.
Fortunately the extended set of mutants used in PIT+ is now integrated in the PIT project\cite{pit+}.

>>>>

\chapter{Mutation testing tools}
\label{ch:mutation-tool}
To cluster mutants we need to select a tool that generate mutants and executes mutants.
Such tools are called mutation testing tools.
Our goal is to find the tool that generates the most amount of mutants. 
\newline
First, we review existing literature. 
Secondly we extend the comparison by adding information about the current state of the tools.
With the combination of the information gathered we make a decision on what tool to use for this thesis.

\section{State of the art documented}
\label{ch:tooling-literature}
The literature used several criteria for comparing mutation testing tools for Java:
\begin{itemize}
  \item Execution time of each tool.
  \item The cost as the number of test cases that needs to be generated and the the number of equivalent mutants that would have to be inspected.
  \item Effectiveness of the mutation adequate test suite of each tool. Every test of the suite adds to the effectiveness score. If a test is removed less than 100\% effectiveness can be achieved. To evaluate the effectiveness of each tool's mutation adequate test suite, a cross-testing technique is applied. The adequate test suite per tool is run on the set of mutants generated by the other tools.
\end{itemize}
Kintis et al. analysed and compared the performance of PIT,  MuJava and Major\cite{Kintis2016AnalysingStudy}. 
They started out with performing a mini literature survey on mutation testing tools used in existing research on test effectiveness in Java. 
With a data set of twelve methods and six Java projects they evaluated each tool using a cross-testing technique.
As a result they found that the most effective mutation adequate test suite was from MuJava, followed by Major and PIT.
The score of the application cost was the inverse of the previous ranking: PIT generated the smallest set of equivalent mutants and required the least test cases.
During their research they also compared the different mutant operators the tools supported. 
An overview of mutant operators(also known as mutators) per tool was created. 
They compared the mutant operators and reported overlapping operators.
\newline
Marki and Lindstrom performed their research on the same three mutation testing tools.\cite{Marki2017MutationJava}. 
The same cross-testing technique used by Kintis et al., was applied for three small Java programs. 
These programs are popular programs used in testing literature.
They found that the mutation tools do not subsume each other. 
"one mutant subsumes a second mutant if every test that kills the first mutant is guaranteed also to kill the second \cite{Ammann2014EstablishingMutants}"
According to the research a ranking of the tools would be as follows; MuJava generated the strongest mutants followed by Major and PIT.
Furthermore they found that MuJava generated significantly more equivalent mutants\cite{Marki2017MutationJava}.
MuJava also had an execution time twice the amount of that of Major and PIT combined.
\newline
Laurent et al., introduced PIT+, a version of PIT with extended set of mutators\cite{Laurent2017AssessingPIT}.
They used the same tests suites that were generated by kintis et al., and combined these into an adequate test suite that would detect the combined set of mutants generated by PIT, MuJava and Major. 
They discovered that the set of mutants generated by PIT+ was stronger than the combined sets generated by the other three tools.
Fortunately the extended set of mutants used in PIT+ is now integrated in the PIT project\cite{pit+}.
